Carrier phase and amplitude estimation for phase shift keying using pilots and data

### 2014-02-04 response.tex

\maketitle

\section*{Reviewer 1}\label{sec:reviewer-1}

\begin{enumerate}
\item\textbf{Comment:} \label{comment:shortersec2}
	This paper is concerned with carrier phase and amplitude estimation for PSK modulated signalling.
The considered least squares estimator seems to be only a minor extension of the one from [7]. Yet, more than 4 pages are dedicated to its derivation. Perhaps this section could be made more concise.
	\\\textbf{Response:}
%\comment{look at shortening sec 2}

In the new manuscript only two pages are dedicated to the description of the algorithm.  An important reason for describing the algorithm is that much of the notation used during the analysis of the estimator is most naturally introduced during the description of the algorithm.  For example, the proof of Lemma~\zref{lem:diffatlambdaL} makes use of the notation $e_i$ and the function $LS(\{d_i, i \in D\})$ introduced during the description of the algorithm.  To qualify this for the reader we have included the following sentences at the beginning of Section~\zref{sec:least-squar-estim}:

### 2014-02-05 response.tex

We have included CRBs for figures in the simulations section.  The follow paragraph has been included:

``Figures~\zref{fig:plotphaseBPSK} and~\zref{fig:plotphaseQPSK} display the non-data aided Cram\'{e}r-Rao lower bound (CRB) for carrier phase given in~\cite{Cowley_crbs_phase_freq_1996}.  To the authors knowledge, simple expression for the CRB for $8$-PSK do not exist in the literature and so the CRB is only shown when $M=2$ (BPSK) and $M=4$ (QPSK).  These CRBs are derived for the non-data aided setting and are only valid for estimators that do not make use of pilots symbols, i.e., $\abs{P}=\emptyset$.''

@@

and in the new introduction we affirm that

``while the specific results given here apply only to the least square estimator, the techniques that we develop to derive these results are now able to be deployed for the analysis of other estimators.''

Thus, the techniques in our paper can potentially be deployed to accurately model the bias and variance of a non-data-aided SNR estimator in order to implement the reviewers proposal.  We confirm, however, that analysing and implementing the reviewers suggested technique is well beyond the scope of the present paper which places primary interest in analysing the least squares estimator.

\item\textbf{Comment:} \label{rev2:biasmse}
	By detailed inspection of Fig.\ 4 showing the jitter variance for amplitude estimates, it is easily observed that $|P| < L$ performs better than $|P| = L$ in the medium-to-low SNR range. In my mind this is somewhat strange, since it should be the other way round (similar to the evolution of the phase curves in Figs.\ 1Ð3 and 5). Can you comment on this phenomenon ?
	\\\textbf{Response:}
In the original manuscript we plotted the \emph{variance} of the amplitude estimator, not the mean square error. That is, the solid lines displayed the asymptotic variance of $\hat{m}_L = \hat{\rho} - \rho_0 G(0)$ and each point from the Monte-Carlo simulations was computed as the unbiased error $\tfrac{1}{T}\sum_{i=1}^T\big(\hat{\rho}_i - \rho_0G(0)\big)^2$.  We chose this approach in order to highlight the correctness of our asymptotic theory. This was described in the 4th paragraph of the old Section~8 containing the simulations section.

We do agree with the reviewer this this plot by itself initially appears misleading.  To alleviate this confusion in the new manuscript we instead plot the average square error computed in the usual way by Monte-Carlo simulation and the solid line is given by the asymptotic squared bias plus variance.  The following paragraph has also included in the simulations section (Section~\zref{sec:simulations}) in the new manuscript.

``Figure~\zref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} 
shows the mean square error of the amplitude estimator when $M=2$ and $L=2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.  The solid lines show the mean square error predicted by the squared bias $\rho_0^2\big(1 - G(0)\big)^2$ plus the variance of $\hat{m}_L = \hat{\rho} - \rho_0G(0)$ predicted by Theorem~\zref{thm:normality}.  That is, the solid lines are given by
\[
\rho_0^2\big(1 - G(0)\big)^2 + \rho_0^2\frac{p B_1 + d B_2}{L}. 
\]
When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.  However, $G(0)$ grows without bound as the variance of the noise increases so the bias is significant when SNR is small.  Figure~\zref{fig:plotampBPSK} suggests our asymptotic theory to be an excellent predictor of the mean square error of the amplitude estimator.
''

\item\textbf{Comment:} \label{rev2:turbosync}
	In a much more powerful approach to the topic addressed in the current paper, the pilot sequence would be employed for a coarse estimation/correction of the carrier phase of the payload data. In the sequel, by embedding carrier phase recovery into an iterative (turbo) decoding scheme\cite{Lottici2004,Noels2005} or any other error correcting method~\cite{580211,974597}, the jitter performance is close to the CRLB for DA estimation even at very small SNRs, which are mainly determined by the convergence region of the (turbo) decoder. By taking into account the fact that modern and powerful communication systems are not conceivable without any FEC scheme, it is not clear to me why I should choose your algorithm in such a context.
	\\\textbf{Response:}
	We have added the following paragraphs to the introduction to place our results in context with exiting turbo synchronisation literature.

``In the last decade, iterative methods for channel estimation that exploit the error correcting code used by the transmitter have appeared~\cite{Lottici2004,Noels2005,Herzet_turbo_synch_proc_IEEE_2007,Herzet_framework_turbo_sync_2007}.  These methods often go by the name of~\emph{turbo synchronisation}~\cite{Herzet_turbo_synch_proc_IEEE_2007}.  
These estimators typically apply the expectation maximisation (EM) algorithm under a Gaussian assumption regarding the noise sequence $\{w_i\}$.  A key problem is that the EM algorithm converges correctly only if initialised with a channel estimate sufficiently close to the true channel parameters.  Methods for efficiently obtaining a close initial channel estimate are still required and because of this the analysis of such methods remains of interest.  The least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation procedure.  However, in some cases, it may be that our estimator is accurate enough, and that the gain obtained by turbo synchronisation will not justify the extra computational cost.

Turbo synchronisation can require a large number of EM iterations.  For applications with low power receivers, such as sensor networks, the complexity of turbo synchronisation might be too great and computationally simple channel estimators are required.  For example, in practice it is common to rely on channel estimates obtained from pilot symbols alone.  In this case the number of pilots symbols must be sufficiently large for channel estimates to be sufficiently accurate.  If more accuracy is required, then more pilots symbols must be used.  Our estimator provides a simple mechanism for increasing accuracy without increasing the number pilot symbols.  Instead, the pilots symbols and a subset of the data symbols can be jointly used for channel estimation.''

@@

%	\comment{Not clear what the reviewer means by ``existing uncoherent methods that do not require estimation of the amplitudes''.}

\item
Some additional comments:
\begin{enumerate}
\item\textbf{Comment:} \label{rev4:shouldbettermotivate}
	The authors should better motivate the interest in using their proposed method in actual communication systems. The lack of motivation for addressing the problem raised in the manuscript is a weak aspect of this work.
	\\\textbf{Response:}
We have expanded the introduction of the paper to better motivate the work.  In particular, the following paragraph has been added:

``That this paper contains novel theoretical developments does not imply that our results are of no immediate practical interest.  The design of our estimator was motivated by a project at the University of South Australia's Institute for Telecommunications Research to develop a narrowband multiuser communication system to collect sensor network data using a satellite in orbit.  This estimator has been implemented on a prototype that was successfully trialled from April to July of 2013 with signals from multiple low power ground-based transmitters being successfully received and decoded by both aircraft and satellites in orbit~\cite{ASRPpromovideo}.  Our theoretical analysis aided design of this system.  Based on our Theorem~\zref{thm:normality} in Section~\zref{sec:stat-prop-least} the performance of the least squares estimator can be accurately predicted without the need for time consuming Monte-Carlo simulations.  This provides easy to use and rigourously justified methods for choosing systems parameters, such as the number and placement of pilot symbols, and also provides useful information to other reciever components, such as channel tracking systems.''

\item\textbf{Comment:} 
	The simulation results should bring comparisons with existing methods that use error correction and turbo decoding techniques, which have shown to perform reasonably well under adverse situations and low SNR levels.
	\\\textbf{Response:}
See our response to comment~\ref{rev2:turbosync} of reviewer~2.

@@

%It can be reasonably expected that the initialiser that is a more accurate estimator of carrier phase and amplitude with also be a better initialiser for a turbo synchronisation proceedure and will thus have a smaller bit error rate. So, for example, we expect our least square estimator to be a beter initialiser at moderate and high SNR and a data-aided approach to be better at low SNR provided there are sufficiently many pilots symbols.

% For this reason that we do not expect simulations of this sort to yield substantially more insight than the mean square error plots that we have presented in the paper.  The primary purpose of our paper is the statistical analysis of the least squares estimator and we feel that

%	\comment{Show PER ASRP ST1 turbo sync results. Comparison with other estimators?}
\item\textbf{Comment:} 
	In Fig.\ 4, the influence of $P$ on the performance is not clear to me and does not seem to be in accordance with other results in the manuscript.
	\\\textbf{Response:} 
Please see response~\ref{rev2:biasmse} to reviewer~2.
%	Please note that Fig.\ 4 in the original submission showed the variance of the amplitude estimator $\hat\rho$ (after bias removal), rather than the %estimator MSE. This was done to validate the theoretical value of the asymptotic variance of $\hat\rho$ predicted by Theorem~2 against Monte-Carlo %simulations. For consistency with the other figures and for easier interpretation, we have replaced the variance plot by a plot of the MSE of $\hat\rho$.
\item\textbf{Comment:} 
	It is important that the authors address an appropriate method for correcting the bias, as they mentioned in the text. A solution to provide the receiver with an estimate of $G(0)$ must be addressed in order to make this contribution applicable in a real communications context.
	\\\textbf{Response:} 
	Biased estimators are quite common in the communications literature and are often used in practice. Prime examples are state-of-the-art signal-to-noise ratio (SNR) estimators for $M$-PSK modulated signals such as those discussed by Pauluzzi~and~Beaulieu~\cite{Pauluzzi2000}. With the exception of the data-aided SNR estimators, all of the estimators in \cite{Pauluzzi2000} are biased and particularly at low SNR, the bias can be large. To the best of the authors' knowledge, the value of the bias of these estimators has only been determined by computer simulation and has not been analysed theoretically. In contrast, the asymptotic analysis of the amplitude estimator $\hat\rho$ presented in Theorem~1 of our manuscript results in a closed-form expression for the bias and thus, lays the foundation for improving the estimator by addressing the bias. The development of methods for removal of the bias is non-trivial and constitutes an entirely new avenue of research.  It is beyond the scope of this manuscript.

\item\textbf{Comment:} \label{R4it2e}
	Some sectioms can be shortened to make the work more readable and concise. This is the case of Section II for instance. Section III is not necessary. It is too short and the main concepts are well known. A reference to a textbook would be sufficient. Section VII is also too short and is not justified.
	\\\textbf{Response:}
The original Section~7 describing the Gaussian noise case has been removed from the revised manuscript.  The original Section~3 has been removed.  Necessary results about circular random variables are now introduced as needed.
%We agree with the reviewer that circularly symmetric random variables (described in Section~3 of the original manuscript) are well known, but point out that  section specifically discussed random variables of the form $Re^{j\Phi}=1+W$, where $W$ is circularly symmetric. The joint pdf of $R$ and $\Phi$ is a crucial ingredient for the asymptotic analysis presented in the manuscript and therefore, we feel that this material must be retained. However, we have modified the way that this material is introduced.  It is now introduced as needed when the asymptotic properties of the least squares estimator are derived.
Regarding Section~2 we refer the reviewer to our response to comment~\eqref{comment:shortersec2} of reviewer~1.

### 2014-02-05 paper.tex
added edits to response. All that's left is to tidy the proof of Lemma 4, and it's useful for me to do this anyway

Aside from this extention of Mackenthun's algorithm, the primary purpose of this paper is to analyse the least squares estimators of carrier phase and amplitude obtained by minimising~\eqref{eq:LSdefn}.  In the literature, analytic study of non-data-aided estimators of carrier phase has primarily involved the derivation of Cramer-Rao lower bounds under the assumption that the noise $\{w_i\}$ is complex Gaussian distributed~\cite{Delmas_exact_crb_psk_2008,950345,Steendam_lowsnr_crb_2001,Cowley_crbs_phase_freq_1996,DAndrea_modified_bounds_1994}.  The bounds themselves are often complicated to compute and analyse and for this reason it is common to make approximations, known as \emph{modified} Cramer-Rao bounds~\cite{DAndrea_modified_bounds_1994,Pollok_mckilliam_crb_cpm_2013}, or to analyse bounds that are valid asymptoticially as the noise variance increases to infinity or decreases to zero~\cite{Moeneclaey_limit_crbs_1998,Steendam_lowsnr_crb_2001,Delmas_exact_crb_psk_2008}.  To the authors knowledge, all of the existing bounds are applicable only in the fully non-data-aided case or in the completely data aided case, that is, when either no pilot symbols exist, or when no data symbols exist.  Deriving bounds for the common practical scenario where there is a mixture of pilot and data symbols might be possible, but would further complicate already complicated bounds.

For the case of carrier amplitude the situations is even more complicated.  A simulation study conducted by Pauluzzi~\cite{Pauluzzi2000} suggests that known non-data-aided estimators for carrier amplitude exhibit bias.  The bias is often small when the noise variance is small, but typically becomes significant as the noise variance increases.  Neverthess, Cramer-Rao bound have been derived for \emph{unbiased} estimators of carrier amplitude~\cite{Alagha_crb_snr_bpsk_qpsk_2001,Delmas_exact_crb_psk_2008}.  Unfortunately, these bounds may not be meaningful since known non-data-aided amplitude estimators appear to be biased~\cite{Pauluzzi2000}.

@@

%Two criticisms might be directed at our asymptotic approach when compared to the Cramer-Rao bounds already developed in the literature.  Perhaps the most obvious is that the results apply asymptotically as $L\to\infty$.  The simulation results we present in Section~\ref{} confirm that our asymptotic results accurately model the behaviour of the least squares estimator for modest numbers of symbols, for example $L=256$

One criticism that might be directed at our asymptotic approach when compared to the Cramer-Rao bounds already developed in the literature is that our specific results apply only to the least square estimator.  One response to this criticism is that the least squares estimator should be a popular practical choice because it can be efficiently computed in $O(L\log L)$ operations and because it is the maximum likelihood estimator under the common assumption that the noise is white and Gaussian.  Perhaps an even more convinving response is that, while the specific results given here apply only to the least squares estimator, the techniques that we develop to derive these results are now able to be deployed for the analysis of other estimators.  For example, in practical systems it is sometimes desirable to place more importance on certain data symbols and use a small constellation size, say BPSK ($M=2$) for these, but to use a larger constellation size, say QPSK ($M=4$), for other less important symbols.  This is related to what is called \emph{unequal error protection}~\cite{Aydinlik_turbo_uep_2008,Sandberg_uep_ldpc_2010}.  It would be reasonably straigthforward to extend our results to allow data and pilot symbols with varying constellation size.  A more involved extension would be the inclusion of carrier frequency as a parameter to be estimated.  The techniques we present, combined with those from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009,McKilliam_mean_dir_est_sq_arc_length2010}, would potentially allow this.  
More ambitiously, our techniques might allow analysis of non-data-aided channel estimators for modern communications systems such as mulitple anntannae systems, systems employing orthogonal frequency division multiplexing, and systems that must widthstand rapidly time varying channels~\cite{4138046,inthft:204506}.  We see the results presented here as a starting point for a potentially potent approach to the analysis of channel estimators that is complementary to the commonplace method of comparing results of Monte-Carlo simulations with Cramer-Rao lower bounds.

@@

In the last decade, iterative methods for channel estimation that exploit the error correcting code used by the transmitter have appeared~\cite{Lottici2004,Noels2005,Herzet_turbo_synch_proc_IEEE_2007,Herzet_framework_turbo_sync_2007}.  These methods often go by the name of~\emph{turbo sychronisation}~\cite{Herzet_turbo_synch_proc_IEEE_2007}.  
These estimators typically apply the expectation maximisation (EM) algorithm under a Gaussian assumption regarding the noise sequence $\{w_i\}$.  A key problem is that the EM algorithm converges correctly only if initialised with a channel estimate sufficiently close to the true channel parameters.  Methods for efficiently obtaining a close initial channel estimate are still required and because of this the analysis of such methods remains of interest.  The least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation proceedure.  However, in some cases, it may be that our estimator is accurate enough, and that the gain obtained by turbo synchronisation will not justify the extra computational cost.

Turbo synchrnoisation can require a large number of EM iterations.  For applications with low power receivers, such as sensor networks, the complexity of turbo sychronisation might be too great and computationally simple channel estimators are required.  For example, in practice it is common to rely on channel estimates obtained from pilot symbols alone.  In this case the number of pilots symbols must be sufficiently large for channel estimates to be sufficiently accurate.  If more accuracy is required, then more pilots symbols must be used.  Our estimator provides a simple mechanism for increasing accuracy without increasing the number pilot symbols.  Instead, the pilots symbols and a subset of the data symbols can be jointly used for channel estimation.  %For computational complexity reasons it may not be desirable to use all of the existing data symbols for channel estimation.  This is easily possible with our estimator by choosing the indices in $D$ to correspond with the subset of the data symbols that are to be used.

@@

@@

Figure~\ref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} 
shows the mean square error of the amplitude estimator when $M=2$ and $L=2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.  The solid lines show the mean square error predicted by the bias $\rho_0\big(1 - G(0)\big)$ squared plus the variance of $\hat{m}_L = \hat{\rho} - \rho_0G(0)$ predicted by Theorem~\ref{thm:normality}.  That is, the solid lines are given by
\[
\rho_0^2\big(1 - G(0)\big)^2 + \rho_0^2\frac{p B_1 + d B_2}{L}. 
\]
When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.  However, $G(0)$ grows without bound as the variance of the noise increases so the bias is significant when SNR is small.  Figure~\ref{fig:plotampBPSK} suggests our asymptotic theory to be an excellent predictor of the mean square error of the amplitude estimator.

Figure~\ref{fig:plotphaseQPSKmultL} shows the MSE of the phase estimator when $M=4$ and $L=32,256, 2048$ and the number of pilots is $\abs{P}=\tfrac{L}{8},L$.  The figure depicts an interesting phenomenon.  When $L=2048$ and $\abs{P} = \tfrac{L}{8} = 256$ the number of pilots symbols is the same as when $L=\abs{P} = 256$.  When the SNR is small (approximately less than \unit[0]{dB}) the least squares estimator using the $256$ pilots symbols and also the $2048-256=1792$ data symbols performs \emph{worse} than the estimator that uses only the $256$ pilots symbols.  A similar phenomenon occurs when $L=256$ and $\abs{P} = \tfrac{L}{8} = 32$.  This behaviour suggests that using the data symbols is counterproductive when the SNR is low.

>>>>

Aside from this extention of Mackenthun's algorithm, the primary purpose of this paper is to analyse the least squares estimators of carrier phase and amplitude obtained by minimising~\eqref{eq:LSdefn}.  In the literature, analytic study of non-data-aided estimators of carrier phase has primarily involved the derivation of Cram\'{e}r-Rao lower bounds under the assumption that the noise $\{w_i\}$ is complex Gaussian distributed~\cite{Delmas_exact_crb_psk_2008,950345,Steendam_lowsnr_crb_2001,Cowley_crbs_phase_freq_1996,DAndrea_modified_bounds_1994}.  The bounds themselves are often complicated to compute and analyse and for this reason it is common to make approximations, known as \emph{modified} Cram\'{e}r-Rao bounds~\cite{DAndrea_modified_bounds_1994,Pollok_mckilliam_crb_cpm_2013}, or to analyse bounds that are valid asymptoticially as the noise variance increases to infinity or decreases to zero~\cite{Moeneclaey_limit_crbs_1998,Steendam_lowsnr_crb_2001,Delmas_exact_crb_psk_2008}.  To the authors knowledge, all of the existing bounds are applicable only in the fully non-data-aided case or in the completely data aided case, that is, when either no pilot symbols exist, or when no data symbols exist.  Deriving bounds for the common practical scenario where there is a mixture of pilot and data symbols might be possible, but would further complicate already complicated bounds.

For the case of carrier amplitude the situations is even more complicated.  A simulation study conducted by Pauluzzi~\cite{Pauluzzi2000} suggests that known non-data-aided estimators for carrier amplitude exhibit bias.  The bias is often small when the noise variance is small, but typically becomes significant as the noise variance increases.  Neverthess, Cram\'{e}r-Rao bound have been derived for \emph{unbiased} estimators of carrier amplitude~\cite{Alagha_crb_snr_bpsk_qpsk_2001,Delmas_exact_crb_psk_2008}.  Unfortunately, these bounds may not be meaningful since known non-data-aided amplitude estimators appear to be biased~\cite{Pauluzzi2000}.

@@

%Two criticisms might be directed at our asymptotic approach when compared to the Cram\'{e}r-Rao bounds already developed in the literature.  Perhaps the most obvious is that the results apply asymptotically as $L\to\infty$.  The simulation results we present in Section~\ref{} confirm that our asymptotic results accurately model the behaviour of the least squares estimator for modest numbers of symbols, for example $L=256$

One criticism that might be directed at our asymptotic approach when compared to the Cram\'{e}r-Rao bounds already developed in the literature is that our specific results apply only to the least square estimator.  One response to this criticism is that the least squares estimator should be a popular practical choice because it can be efficiently computed in $O(L\log L)$ operations and because it is the maximum likelihood estimator under the common assumption that the noise is white and Gaussian.  Perhaps an even more convinving response is that, while the specific results given here apply only to the least squares estimator, the techniques that we develop to derive these results are now able to be deployed for the analysis of other estimators.  For example, in practical systems it is sometimes desirable to place more importance on certain data symbols and use a small constellation size, say BPSK ($M=2$) for these, but to use a larger constellation size, say QPSK ($M=4$), for other less important symbols.  This is related to what is called \emph{unequal error protection}~\cite{Aydinlik_turbo_uep_2008,Sandberg_uep_ldpc_2010}.  It would be reasonably straigthforward to extend our results to allow data and pilot symbols with varying constellation size.  A more involved extension would be the inclusion of carrier frequency as a parameter to be estimated.  The techniques we present, combined with those from~\cite{McKilliamFrequencyEstimationByPhaseUnwrapping2009,McKilliam_mean_dir_est_sq_arc_length2010}, would potentially allow this.  
More ambitiously, our techniques might allow analysis of non-data-aided channel estimators for modern communications systems such as mulitple anntannae systems, systems employing orthogonal frequency division multiplexing, and systems that must widthstand rapidly time varying channels~\cite{4138046,inthft:204506}.  We see the results presented here as a starting point for a potentially potent approach to the analysis of channel estimators that is complementary to the commonplace method of comparing results of Monte-Carlo simulations with Cram\'{e}r-Rao lower bounds.

@@

In the last decade, iterative methods for channel estimation that exploit the error correcting code used by the transmitter have appeared~\cite{Lottici2004,Noels2005,Herzet_turbo_synch_proc_IEEE_2007,Herzet_framework_turbo_sync_2007}.  These methods often go by the name of~\emph{turbo synchronisation}~\cite{Herzet_turbo_synch_proc_IEEE_2007}.  
These estimators typically apply the expectation maximisation (EM) algorithm under a Gaussian assumption regarding the noise sequence $\{w_i\}$.  A key problem is that the EM algorithm converges correctly only if initialised with a channel estimate sufficiently close to the true channel parameters.  Methods for efficiently obtaining a close initial channel estimate are still required and because of this the analysis of such methods remains of interest.  The least squares estimator we describe here is a good candidate as an initialiser for a turbo synchronisation procedure.  However, in some cases, it may be that our estimator is accurate enough, and that the gain obtained by turbo synchronisation will not justify the extra computational cost.

Turbo synchronisation can require a large number of EM iterations.  For applications with low power receivers, such as sensor networks, the complexity of turbo synchronisation might be too great and computationally simple channel estimators are required.  For example, in practice it is common to rely on channel estimates obtained from pilot symbols alone.  In this case the number of pilots symbols must be sufficiently large for channel estimates to be sufficiently accurate.  If more accuracy is required, then more pilots symbols must be used.  Our estimator provides a simple mechanism for increasing accuracy without increasing the number pilot symbols.  Instead, the pilots symbols and a subset of the data symbols can be jointly used for channel estimation.  %For computational complexity reasons it may not be desirable to use all of the existing data symbols for channel estimation.  This is easily possible with our estimator by choosing the indices in $D$ to correspond with the subset of the data symbols that are to be used.

@@

Figures~\ref{fig:plotphaseBPSK} and~\ref{fig:plotphaseQPSK} display the non-data aided Cram\'{e}r-Rao lower bound (CRB) for carrier phase given in~\cite{Cowley_crbs_phase_freq_1996}.  To the authors knowledge, simple expression for the CRB for $8$-PSK do not exist in the literature and so the CRB is only shown when $M=2$ (BPSK) and $M=4$ (QPSK).  These CRBs are derived for the non-data aided setting and are only valid for estimators that do not make use of pilots symbols, i.e., $\abs{P}=\emptyset$.

@@

Figure~\ref{fig:plotampBPSK} %,~\ref{fig:plotampQPSK}~and~\ref{fig:plotamp8PSK} 
shows the mean square error of the amplitude estimator when $M=2$ and $L=2048$ and when the number of pilots symbols is $\abs{P} = 0, \tfrac{L}{2}, L$.  The solid lines show the mean square error predicted by the squared bias $\rho_0^2\big(1 - G(0)\big)^2$ plus the variance of $\hat{m}_L = \hat{\rho} - \rho_0G(0)$ predicted by Theorem~\ref{thm:normality}.  That is, the solid lines are given by
\[
\rho_0^2\big(1 - G(0)\big)^2 + \rho_0^2\frac{p B_1 + d B_2}{L}. 
\]
When SNR is large $G(0)$ is close to $1$ and the bias of the amplitude estimator is small.  However, $G(0)$ grows without bound as the variance of the noise increases so the bias is significant when SNR is small.  Figure~\ref{fig:plotampBPSK} suggests our asymptotic theory to be an excellent predictor of the mean square error of the amplitude estimator.

Figure~\ref{fig:plotphaseQPSKmultL} shows the MSE of the phase estimator when $M=4$ and $L=32,256, 2048$ and the number of pilots is $\abs{P}=\tfrac{L}{8},L$.  The figure depicts an interesting phenomenon.  When $L=2048$ and $\abs{P} = \tfrac{L}{8} = 256$ the number of pilots symbols is the same as when $L=\abs{P} = 256$.  When the SNR is small (approximately less than \unit[0]{dB}) the least squares estimator using the $256$ pilots symbols and also the $2048-256=1792$ data symbols performs \emph{worse} than the estimator that uses only the $256$ pilots symbols.  A similar phenomenon occurs when $L=256$ and $\abs{P} = \tfrac{L}{8} = 32$.  This behaviour suggests that using the data symbols is counterproductive when the SNR is low and the proportion of pilot symbols is sufficiently large.
